---
title: "STAT 170 Final Project Info"
output:
  word_document: default
  html_document: default
date: "2023-11-15"
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

# Setting up data

```{r}
insurance <- read.csv("~/Desktop/UC_Riverside/2023-2024/Fall/STAT 170/Final Project/insurance.csv")
# View(insurance)
```

```{r}
library(tidyverse)
insurance
```

Setting variables in R

```{r}
age <- insurance$age
sex <- insurance$sex
bmi <- insurance$bmi
children <- insurance$children
smoker <- insurance$smoker
region <- insurance$region
charges <- insurance$charges
```

# Plots of Data
Numerical Data: Histograms

```{r}
par(mfrow = c(2,2))
hist1 <- hist(age)
hist2 <- hist(bmi)
hist3 <- hist(children)
hist4 <- hist(charges)
```

Categorical Data: Frequency tables
```{r}
par(mfrow = c(3,1))
table1 <- table(sex)
table2 <- table(smoker)
table3 <- table(region)

barplot(table1)
barplot(table2)
barplot(table3)
```

Box Plots

```{r}
par(mfrow = c(2,2))
boxplot(age, main = "Box plot of age", horizontal = T)
boxplot(bmi, main = "Box plot of bmi", horizontal = T)
boxplot(children, main = "Box plot of children", horizontal = T)
boxplot(charges, main = "Box plot of charges", horizontal = T)
```


# Summary Statistics

```{r}
print("Age Summary")
print(summary(age))
print("BMI Summary")
print(summary(bmi))
print("Children Summary")
print(summary(children))
print("Charges Summary")
print(summary(charges))
```


# Pairwise Scatterplots + Coefficients of Correlation for all variables (Multicollinearity)
```{r}
library("ggplot2")
library("gridExtra")

library("GGally")
ggpairs(insurance) + theme()
```

Multicollinearity Correlation Values for predictors

```{r}
# age, bmi, children, charges
print("Correlation Coefficient of age ~ charges")
print(cor(age, charges))
print("Correlation Coefficient of bmi ~ charges")
print(cor(bmi, charges))
print("Correlation Coefficient of children ~ charges")
print(cor(children, charges))
```


# Multicollinearity Summary of Linear Model

```{r}
ins.mlr <- lm(charges~age+sex+bmi+children+smoker+region, data = insurance)
summary(ins.mlr)
```

# Variance Inflation Factor (VIF)

```{r}
# Using car Library
# Used in Analysis
library(car)
vif(ins.mlr)
```

```{r}
# Using rms library
library(rms)
vif(ins.mlr)
```



# ANALYSIS 2

## Transform the `charges` variable

```{r}
# Log-transformed to make Charges variable more normal distributed
charges_log <- log(charges)
hist(charges_log)
boxplot(charges_log, main = "Box plot of charges", horizontal = T)
```

# Residual Analysis for First-Order Predictors
## Residuals vs. Fitted Plot

```{r}
# plot(log_model)

# ANOVA
# anova(log_model)
```


# $R^2$ Values for insurance dataset with log-transformed `Charges` data:

```{r}
library(leaps)
insurance_all <- regsubsets(charges_log ~ age+sex+bmi+children+smoker+region, data = insurance,
                            nbest = 1, nvmax = 5)
summary(insurance_all)$rsq # R^2
summary(insurance_all)$adjr2 # Adjusted R^2
```




# Full Model Selection (Second Order Terms)

```{r}
library(leaps)
charges_all <- regsubsets(charges_log ~.^2 + I(age^2) + I(bmi^2) + I(children^2), data = insurance, nbest = 1, nvmax = 5)
summary(charges_all)$which
summary(charges_all)$adjr2
```

# Step Model Selection

```{r}
full_model <- lm(charges_log ~.^2 + I(age^2) + I(bmi^2) + I(children^2), data = insurance)
null_model <- lm(charges_log ~ 1, data = insurance)
step_model1 <- step(null_model, scope = list(lower = null_model, upper = full_model), direction = "both",test="F")
summary(step_model1)

```


# Multicollinearity Analysis (on chosen model)

```{r}

```

# Correlation Coefficients

```{r}
# age, bmi, children, charges_log
print("Correlation Coefficient of age ~ charges_log")
print(cor(age, charges_log))
print("Correlation Coefficient of bmi ~ charges_log")
print(cor(bmi, charges_log))
print("Correlation Coefficient of children ~ charges_log")
print(cor(children, charges_log))
```



# Step-wise Regression

```{r}
all_possible_data <- select(insurance, -charges)
charges_log <- log(insurance$charges)
full_model <- lm(charges_log ~.^2 + I(age^2) + I(bmi^2) + I(children^2), data = all_possible_data)
null_model <- lm(charges_log ~ 1, data = insurance)
step_model1 <- step(null_model, scope = list(lower = null_model, upper = full_model), direction = "both",test="F")
```


# Multicollinearity Summary

```{r}
log_model <- lm(charges_log ~ age+sex+bmi+children+smoker+region, data = insurance)
summary(log_model)
```


# VIF on Chosen Model

```{r}
# Using car Library
vif(step_model1)
```
